---
title: "Canada glaciers trend analisys"
author: "Stanislav"
output: 
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    keep_tex: yes
bibliography: bibl.bib
urlcolor: blue
params:
  filename: glacier_data.csv
linkcolor: magenta
geometry: margin=0.8in
header-includes:
   - \usepackage[T2A]{fontenc}
   - \usepackage[utf8]{inputenc}
   - \usepackage[russian]{babel}
   - \usepackage{caption}
   - \usepackage{subcaption}
---

```{r setup, include=FALSE}
library(tidyr)
library(ggplot2)
library(ggvis)
library(dplyr)
options(scipen = 0, digits = 3)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

```

## Inroduction
```{r, include=FALSE}

setwd("~/RProjects/R_courses/Statistica/Part2")
glacier <- read.csv(params$filename, na.strings = "..")
```
The data used in study is taken from:
<https://open.canada.ca>[^1]. The algorithm is taken from @AntonovErmakov_RandomCubaturesQMC.

[^1]: [Goverment of Canada site](https://open.canada.ca/data/en/dataset/aca5e1de-c234-5372-b714-ae20300fb6ce)


This data set contains `r nrow(glacier)` measurments of `r nlevels(glacier$GEO)` Canadian glaciers mass balance, collected from `r min(glacier$REF_DATE)` till `r max(glacier$REF_DATE)`. Namely, the file includes these glaciers:
```{r}
levels(glacier$GEO)
```

## Hypothesis

We are interested in finding out whether there is a statistically significant change in mass balance over the observed time period.
purposes we use **R** (version `r getRversion()`) and an appropriate statistical test called _t-test_: $$t = \frac{\overline{x} - \mu_0}{s/\sqrt{n]}}.$$

The workflow is as follows:

1. >Read file
2. [text](Run t-test for each glacier and collect p-values)
3. *Support the evidence with*
 + **a table of results;**
 + a plot the could help demonstrate the effect.
 
```{r, cache=TRUE}

glacier <- glacier %>% 
  select(REF_DATE, GEO, VALUE, Type.of.measure) %>% 
  filter(Type.of.measure == 'Annual mass balance') %>% 
  separate(GEO, c('Name', 'Location'), sep = ' - ')

# descriptive analisys
g1 <- glacier %>% 
  group_by(Name) %>% 
  summarise(YearsObserved = n(),
            MeanChange = mean(VALUE, na.rm = T),
            WorstChange = min(VALUE, na.rm = T),
            WorstYear = REF_DATE[which.min(VALUE)])


#t-test

g2 <- glacier %>% 
  group_by(Name) %>% 
  do({
    tt <- t.test(.$VALUE, alternative = 'less', mu = 0, conf.level = 0.99)
    tibble(Pvalue = tt$p.value,
               ConfidentLimit = tt$conf.int[2])
  })
left_join(g1, g2, by = 'Name') %>% 
  knitr::kable(caption = 'descriptive statistics', digits = c(0,0,2,0,0,10,2))

```
 
```{r}
#ggplot

# ggplot(glacier, aes(REF_DATE, VALUE))+
#   geom_line()+
#   geom_hline(data = g1, aes(yintercept = MeanChange), color = 'coral',
#              lty = 'dashed', alpha = 0.8)+
#   facet_wrap(~Name)
ggvis(glacier, ~REF_DATE, ~VALUE, stroke = ~Name) %>% 
  layer_lines()

```

## Analisys
<u> првиет </u>

## Bibliography

 
 